---
title: "Enrichment Analysis"
output: html_document
---

```{r Read in}
sig.bcr <- readRDS("/misc/pichu/data/lara/objects/significantGenes_bcr.rds")
sig.kmt <- readRDS("/misc/pichu/data/lara/objects/significantGenes_kmt.rds")

upGenes.bcr <- readRDS("/misc/pichu/data/lara/objects/upregulatedGenes_bcr.rds")
downGenes.bcr<- readRDS("/misc/pichu/data/lara/objects/downregulatedGenes_bcr.rds")
upGenes.kmt <- readRDS("/misc/pichu/data/lara/objects/upregulatedGenes_kmt.rds")
downGenes.kmt<- readRDS("/misc/pichu/data/lara/objects/downregulatedGenes_kmt.rds")
```

```{r}
ids.bcr <- data.frame(ENSEMBL = rownames(sig.bcr))
ids.kmt <- data.frame(ENSEMBL = rownames(sig.kmt))

```

# GO Enrichment Analysis


```{r project enrichment analysis}


ego.bcr <- enrichGO(gene     = ids.bcr$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "BP", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)

ego.kmt <- enrichGO(gene     = ids.kmt$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "BP", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)


# IDs of filtered log2foldchange Genes
ids.poslog.bcr <- data.frame(ENSEMBL = rownames(upGenes.bcr))
ids.neglog.bcr <- data.frame(ENSEMBL = rownames(downGenes.bcr))
ids.poslog.kmt <- data.frame(ENSEMBL = rownames(upGenes.kmt))
ids.neglog.kmt <- data.frame(ENSEMBL = rownames(downGenes.kmt))

ego.poslog.bcr <- enrichGO(gene     = ids.poslog.bcr$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "BP", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)

ego.poslog.kmt <- enrichGO(gene     = ids.poslog.kmt$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "BP", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)
# doesn't work
ego.neglog.bcr <- enrichGO(gene     = ids.neglog.bcr$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "BP", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)

# doesn't work 
ego.neglog.kmt <- enrichGO(gene     = ids.neglog.kmt$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "BP", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)

#Molecular function--------------------------------------------------------
ego.bcr2 <- enrichGO(gene     = ids.bcr$ENSEMBL,
                  OrgDb         = org.Hs.eg.db,
                  keyType       = 'ENSEMBL',
                  ont           = "MF", 
                  pAdjustMethod = "BH",
                  pvalueCutoff  = 0.01,
                  qvalueCutoff  = 0.05,
                  readable = TRUE)

cnetplot(ego.bcr2)
ggsave("/misc/pichu/data/lara/figures/CNETPlot_MF_BCR.pdf", width = 8, height = 5)
```
# Plots GO Enrichment Analysis

```{r project plot enrichtment analysis}
# For downregulated genes no results

dotplot(ego.poslog.bcr)
goplot(ego.poslog.bcr)
cnetplot(ego.poslog.bcr) +
  ggplot2::ggtitle("GO Enrichment Analysis for Upregulated Genes in BCR") +
  theme(
   plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
   legend.text = element_text(size = 15),
   legend.title = element_text(size = 13),
  )
ggsave("/misc/pichu/data/lara/figures/CNETPlot_upregulated_BCR.pdf", width = 6, height = 5)

cnetplot(ego.poslog.kmt) +
  ggplot2::ggtitle("GO Enrichment Analysis for Upregulated Genes in KMT") +
  theme(
   plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
   legend.text = element_text(size = 15),
   legend.title = element_text(size = 13),
  )
ggsave("/misc/pichu/data/lara/figures/CNETPlot_upregulated_KMT.pdf", width = 10, height = 8)

```

# Reactome Enrichment Analysis with up- and downregulated Genes

```{r project reactome enrichment analysis}

#Ensembl IDs to Entrezid IDs from the significant genes----

# Filter and add Column name

valid_ids <- keys(org.Hs.eg.db, keytype = "ENSEMBL")

poslogFiltered.bcr <- poslog.bcr[rownames(poslog.bcr) %in% valid_ids,] %>%
  tibble::rownames_to_column("ENSEMBL")

poslogFiltered.kmt <- poslog.kmt[rownames(poslog.kmt) %in% valid_ids,] %>%
  tibble::rownames_to_column("ENSEMBL")

neglogFiltered.bcr <- neglog.bcr[rownames(neglog.bcr) %in% valid_ids,] %>%
  tibble::rownames_to_column("ENSEMBL")

neglogFiltered.kmt <- neglog.kmt[rownames(neglog.kmt) %in% valid_ids,] %>%
  tibble::rownames_to_column("ENSEMBL")

# Check

poslogFiltered.bcr$ENSEMBL %in% keys(org.Hs.eg.db, keytype = "ENSEMBL")
poslogFiltered.kmt$ENSEMBL %in% keys(org.Hs.eg.db, keytype = "ENSEMBL")
neglogFiltered.bcr$ENSEMBL %in% keys(org.Hs.eg.db, keytype = "ENSEMBL")
neglogFiltered.kmt$ENSEMBL %in% keys(org.Hs.eg.db, keytype = "ENSEMBL")

# Biological Translator

entrezid.poslog.bcr<- bitr(poslogFiltered.bcr$ENSEMBL, fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)

entrezid.poslog.kmt<- bitr(poslogFiltered.kmt$ENSEMBL, fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)

entrezid.neglog.bcr<- bitr(neglogFiltered.bcr$ENSEMBL, fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)

entrezid.neglog.kmt<- bitr(neglogFiltered.kmt$ENSEMBL, fromType = "ENSEMBL",
        toType = "ENTREZID",
        OrgDb = org.Hs.eg.db)

# Duplicates in the ENS-IDs in KMT Samples
entrezid.poslog.bcr$log2FoldChange <- poslogFiltered.bcr$log2FoldChange[match(entrezid.poslog.bcr$ENSEMBL, poslogFiltered.bcr$ENSEMBL)]
entrezid.poslog.kmt$log2FoldChange <- poslogFiltered.kmt$log2FoldChange[match(entrezid.poslog.kmt$ENSEMBL, poslogFiltered.kmt$ENSEMBL)] 
entrezid.neglog.bcr$log2FoldChange <- neglogFiltered.bcr$log2FoldChange[match(entrezid.neglog.bcr$ENSEMBL, neglogFiltered.bcr$ENSEMBL)] 
entrezid.neglog.kmt$log2FoldChange <- neglogFiltered.kmt$log2FoldChange[match(entrezid.neglog.kmt$ENSEMBL, neglogFiltered.kmt$ENSEMBL)] 

#Create geneList with Entrezids = log2foldchange----

geneList.poslog.bcr <- setNames(
  entrezid.poslog.bcr$log2FoldChange,
  entrezid.poslog.bcr$ENTREZID
)

geneList.poslog.kmt <- setNames(
  entrezid.poslog.kmt$log2FoldChange,
  entrezid.poslog.kmt$ENTREZID
)

geneList.neglog.bcr <- setNames(
  entrezid.neglog.bcr$log2FoldChange,
  entrezid.neglog.bcr$ENTREZID
)

geneList.neglog.kmt <- setNames(
  entrezid.neglog.kmt$log2FoldChange,
  entrezid.neglog.kmt$ENTREZID
)

#Reactome enrichment analysis----

data(geneList.poslog.bcr, package="DOSE")
de.poslog.bcr <- names(geneList.poslog.bcr)[abs(geneList.poslog.bcr) > 0.05]
rea.poslog.bcr <- enrichPathway(gene=de.poslog.bcr, pvalueCutoff = 0.05, readable=TRUE)

data(geneList.poslog.kmt, package="DOSE")
de.poslog.kmt <- names(geneList.poslog.kmt)[abs(geneList.poslog.kmt) > 0.05]
rea.poslog.kmt <- enrichPathway(gene=de.poslog.kmt, pvalueCutoff = 0.05, readable=TRUE)

data(geneList.neglog.bcr, package="DOSE")
de.neglog.bcr <- names(geneList.neglog.bcr)[abs(geneList.neglog.bcr) > 0.05]
rea.neglog.bcr <- enrichPathway(gene=de.neglog.bcr, pvalueCutoff = 0.05, readable=TRUE)

data(geneList.neglog.kmt, package="DOSE")
de.neglog.kmt <- names(geneList.neglog.kmt)[abs(geneList.neglog.kmt) > 0.05]
rea.neglog.kmt <- enrichPathway(gene=de.neglog.kmt, pvalueCutoff = 0.05, readable=TRUE)


```

# Plots Reactome Enrichment Analysis

```{r project plots reactome enrichment analysis}

barplot(rea.poslog.bcr, showCategory = 10, title = "Top Enriched Reactome Pathways for Upregulated Genes in BCR")
barplot(rea.poslog.kmt, showCategory = 10, title = "Top Enriched Reactome Pathways for Upregulated Genes in KMT")
barplot(rea.neglog.bcr, showCategory = 10, title = "Top Enriched Reactome Pathways for Downregulated Genes in BCR")
barplot(rea.neglog.kmt, showCategory = 10, title = "Top Enriched Reactome Pathways for Downregulated Genes in KMT") # length zero

cnetplot(rea.poslog.bcr, showCategory = 5, circular = TRUE, colorEdge = TRUE) + ggplot2::ggtitle("Top Enriched Reactome Pathways for Upregulated Genes in BCR")
cnetplot(rea.poslog.kmt, showCategory = 5, circular = TRUE, colorEdge = TRUE) + ggplot2::ggtitle("Top Enriched Reactome Pathways for Upregulated Genes in KMT")
cnetplot(rea.neglog.bcr, showCategory = 5, circular = TRUE, colorEdge = TRUE) + ggplot2::ggtitle("Top Enriched Reactome Pathways for Downregulated Genes in BCR")
cnetplot(rea.neglog.kmt, showCategory = 5, circular = TRUE, colorEdge = TRUE) + ggplot2::ggtitle("Top Enriched Reactome Pathways for Downregulated Genes in KMT") # zero

```

# Overlap between diff. expressed Genes in BCR and KMT

```{r project find overlaps}

poslog.bcr <- tibble::rownames_to_column(poslog.bcr,"ENSEMBL")
poslog.kmt <- tibble::rownames_to_column(poslog.kmt,"ENSEMBL")
neglog.bcr <- tibble::rownames_to_column(neglog.bcr,"ENSEMBL")
neglog.kmt <- tibble::rownames_to_column(neglog.kmt,"ENSEMBL")

overlap_upregulated_genes <- intersect(poslog.bcr$ENSEMBL, poslog.kmt$ENSEMBL)
overlap_downregulated_genes <- intersect(neglog.bcr$ENSEMBL, neglog.kmt$ENSEMBL)

length(overlap_upregulated_genes) # = 3. "ENSG00000231419" "ENSG00000182704" "ENSG00000173918"
length(overlap_downregulated_genes) # = 7. "ENSG00000210082" "ENSG00000207008" "ENSG00000207392" "ENSG00000234093" "ENSG00000231369" "ENSG00000213939" "ENSG00000227034"

# Get IDs from upregulated Genes as Input for STRING 

cat(poslog.kmt$ENSEMBL, sep = "\n")
cat(poslog.kmt$ENSEMBL, sep = "\n")
```

# Change log2foldchange and padj for KMT, so it fits for STRING 

```{r project change log and padj for KMT}
padjChanged.kmt <- resOrdered.kmt %>%
  dplyr::filter(padj < 0.05)

poslogChanged.kmt <- padjChanged.kmt %>%
  dplyr::filter(log2FoldChange > 4.8)
neglogChanged.kmt <- padjChanged.kmt %>%
  dplyr::filter(log2FoldChange < -0.5)

poslogChanged.kmt <- tibble::rownames_to_column(poslogChanged.kmt,"ENSEMBL")
cat(poslogChanged.kmt$ENSEMBL, sep = "\n")

# Get IDs from genes as Input for String
```

# Overlap between BCR and KMT with one DDS-object

```{r project one dds object for kmt and bcr to study overlaps}

# Create meta data table with BCR and KMT as well as diagnosis and relapsed
metaBoth <- rbind(meta.bcr, meta.kmt)

# Create counts table
countsBoth <- merge(counts.bcr, counts.kmt, by = "row.names", all = TRUE)
rownames(countsBoth) <- countsBoth$Row.names
countsBoth <- countsBoth[,-1]

# Create DDS Object
ddsBoth <- DESeqDataSetFromMatrix(countData = countsBoth,
                                  colData = metaBoth,
                                  design = ~ condition + Relapsed)
# Prefiltering low counts
smallestGroupSize <- 2
keep <- rowSums(counts(ddsBoth) >= 10) >= smallestGroupSize
ddsBoth <- ddsBoth[keep,]

# DeSeq2
ddsBoth <- DESeq(ddsBoth)
saveRDS(ddsBoth, file = "/misc/pichu/data/lara/objects/dds/dds_kmt_bcr.rds")
resBoth <- results(ddsBoth)

# How to get information about relapsed vs diagnosis

# Order p-values
resBothOrdered <- resBoth[order(resBoth$pvalue),]

# Get Summary
summary(resBoth)

# Create Data Frame
resBothOrdered.df <- as.data.frame(resBothOrdered)

# Plot
plotCounts(ddsBoth, gene=which.min(resBoth.df$padj), intgroup="Relapsed")

# Filtering
padjBoth <- resBothOrdered.df %>%
  dplyr::filter(padj < 0.05)
saveRDS(padjBoth, file = "/misc/pichu/data/lara/objects/significantGenes_kmt&bcr.rds")

poslogBoth <- padjBoth %>%
  dplyr::filter(log2FoldChange > 5)

neglogBoth <- padjBoth %>%
  dplyr::filter(log2FoldChange < -5)

# Get upregulated and downregulated Gene IDs 
poslogBoth <- tibble::rownames_to_column(poslogBoth,"ENSEMBL")
neglogBoth <- tibble::rownames_to_column(neglogBoth,"ENSEMBL")

cat(poslogBoth$ENSEMBL, sep = "\n") # used for STRING
length(poslogBoth$ENSEMBL)

cat(neglogBoth$ENSEMBL, sep = "\n")
length(neglogBoth$ENSEMBL)
